---
title: "Data 1901 Assignment"
authors: Blake, Leo, Ari, August Gonzales, Harry Le
format:
    html:
        embed-resources: true
editor: visual
---

STILL NEED TO EMBED DATA PROPERLY INTO THE DOCUMENT(OMIT ANY DATA, TURN AGE INTO QUALATATIVE, AND USE A THEME) Also might wanna use Levenes test to show only weak correlations

```{r setup, echo=FALSE, warning=FALSE, message=FALSE, include = FALSE}
library(tidyverse)
library(ggplot2)
library(plotly)
library(readxl)
library(dplyr)

file_path <- "all_data.xlsx"
all_data <- read_excel("all_data.xlsx", sheet = "pooled123")
library(car)
all_data$anxiety_change <- all_data$p_vra - all_data$baseline_vra
leveneTest(anxiety_change ~ choice, data = all_data)
```

## Executive Summary

Our analysis within this report found that typically no anxiety change relationship could be found between enviroments. Although, notably the range for change in anxiety for those with a sunny environment was much greater than those with a snowy environment, indicating a small correlation. Moreover, the analysis conveyed that the spreads of the data decreased for both environments when given the choice to choose. As such the environment the user experiences has little association with the change in anxiety, although when given a choice a more decisive result may occur although the data is not strong enough to confirm this.

## Initial Data Analysis

**Overview**

The data was sourced through an experimental survey examining 337 partiicipants for the VR nocebo affect. Our investigation has a unique focus on age groups within the experiments, and how a VR environment may cause a impact on the fluctuation of the participants reported anxiety levels. With these variables we had to reclassify age from a quantitative continuous variable into a qualitative nominal to utilize groups for which the anxiety changed, while the rest remained. These being the VR environment as a qualitative nominal, and the anxiety level's quantitative discrete variable, and choice as a qualitative nominal.

```{r} # I think this is correct for the graph, for the change in anxiety I dunno. It also shows that a sunny environment may have not as large of an affect in anxiety compared to snowy, Although notable the Simpsons affect might take affect here as the min and max for sunny is normally higher which means that their increase or decrease in anxiety might make it seem like theres not as big of an affect. (Median is always equal to or less than) ALSO NOTE: This may not be the best way, as we only know the change E.g Final - Intial, ALSO THIS IS CHANGE, LOWER ONLY MEANS A LOWER CHANGE NOT REDUCED ANXIETY.}

all_data <- all_data %>%
  mutate(change_in_anxiety = p_vra - baseline_vra,  # Calculate the change in anxiety
         age_group = cut(age, 
                          breaks = c(15, 20, 25, 30, 35, 40), 
                          labels = c("15-20", "20-25", "25-30", "30-35", "35-40"),
                          right = FALSE))  # Create the age groups

# Filter out outliers for change in anxiety (below 0 or above 5)
all_data_clean <- all_data %>%
  filter(change_in_anxiety >= -4 & change_in_anxiety <= 10)  # Remove rows where change in anxiety is below 0 or above 5

# Create the boxplot with the desired y-axis scale using the change in anxiety
p <- ggplot(all_data_clean, aes(x = env, y = change_in_anxiety, fill = env)) +
  geom_boxplot(na.rm = TRUE) +  # Ensure that missing values don't break the plot
  facet_wrap(~ age_group) +  # Separate by age groups
  labs(title = "Change in Anxiety Levels by System Environment and Age Group",
       x = "System Environment",
       y = "Change in Anxiety Level",
       fill = "Environment") +
  theme_minimal() +
  scale_y_continuous(limits = c(-5, 10), 
                     breaks = seq(-4, 10, by = 2))  # Y-axis ticks every 2 units

# Convert the plot to an interactive plot
ggplotly(p)
```

```{r, fig.height=30, fig.width=20}
# This graph shows that when given a choice, the max and mins become a lot closer to the quatiles, hence less outliers in the data. Although it is notable that the median change of anxiety does not change that much so they still experience similar amounts of change. NOTE: LOOK AT HTML FOR FULL VIEW

all_data <- all_data %>%
  mutate(change_in_anxiety = p_vra - baseline_vra,  # Calculate the change in anxiety
         age_group = cut(age, 
                          breaks = c(15, 20, 25, 30, 35, 40), 
                          labels = c("15-20", "20-25", "25-30", "30-35", "35-40"),
                          right = FALSE))  # Create the age groups

# Filter out outliers for change in anxiety (below 0 or above 5)
all_data_clean <- all_data %>%
  filter(change_in_anxiety >= -4 & change_in_anxiety <= 10)  # Remove rows where change in anxiety is below 0 or above 5


# Now, create the faceted boxplot with the two sections:
# One for participants who chose their environment, 
# and one for participants who did not choose their environment

p <- ggplot(all_data_clean, aes(x = env, y = change_in_anxiety, fill = env)) +
  geom_boxplot(na.rm = TRUE) +  # Ensure that missing values don't break the plot
  facet_wrap(~ choice + age_group, ncol = 1) +  # Facet by choice condition and age group
  labs(title = "Change in Anxiety Levels by System Environment",
       x = "System Environment",
       y = "Change in Anxiety Level",
       fill = "Environment") +
  theme_minimal() +
  scale_y_continuous(limits = c(-10, 10), 
                     breaks = seq(-5, 10, by = 5)) +  # Y-axis ticks every 2 units
  theme(strip.text = element_text(size = 10),  # Adjust facet label size
        strip.background = element_blank(),  # Remove background color from facet labels
        panel.spacing = unit(2, "lines"),  # Increase spacing between panels
        axis.text.y = element_text(size = 12, angle = 0, hjust = 1),  # Increase size and adjust angle of y-axis labels
        axis.title.y = element_text(size = 14),  # Increase the size of y-axis title
        axis.text.x = element_text(size = 10))  # Adjust x-axis label size

# Convert the plot to an interactive plot and increase vertical size with margin adjustments
ggplotly(p) %>%
  layout(height = 1000,  # Adjust the height further
         margin = list(l = 100, r = 100, t = 100, b = 100)) 
```

**Limitations**

This study is limited by its sample size, this entails that the experiment will not be able to get perfectly accurate results. Furthermore, external factors like mood and behaviors before entering the VR simulation will not be able to be controlled, as such the data could vary depending on the sets of age groups.

**Assumption**

Within this report there was an assumption that people were honest with their symptoms, and variation in anxiety levels. Along with having any non permitted knowledge on how the test is conducted beyond what has been told. Furthermore, it is assumed that neither the conductors or participants had any bias on the data. It is also assumed the test were completed in the same conditions, such that no co-founding variables would have altered the data sets. Additionally, data cleaning had to take place to remove outliers, these were anxiety levels that were outside of the possible range, as these were assumed to be a mistake. Furthermore, the data was concluded down to the age groups between 15 and 40 as these had the most data, and were deemed the most representative of real world scenarios.

## Research Question

GRAPHS: CHOICE/NO CHOICE, (AGE, ENVIROMENT, ANXIETY IN ONE), INITIAL AND FINAL ANXIETY(so if large at beginning did it reduce), ANYTHING THAT MAY RELATE

Explores a research question, using numerical and/or graphical summaries (i.e., ggplot, gganimation, maps) which give insight.

**Q) Does the volition on the type of VR environment impact the severity of anxiety level deviation**

The data conveyed that the change in anxiety levels between users of sunny, and snowy environments varied by minuscule amounts. This is shown in the median, mean, and quartiles of the data. The average median among all age groups for sunny environments was a change of 1.4, while snowy was 1.6, revealing that snowy environments had a slightly higher change. Furthermore, the average range of IQR was 3.6 for sunny environments, and 3.1 for snowy. This implies that snowy environments more narrow IQR may be a resultant of a slightly more decisive, consistent and reduced overall change. Hence, the small difference in median, and IQR range indicate that system environment has little to no correlation with anxiety level changes. Furthermore, the data implied that little change would result from a choice, as the average median for participants who chose was, 1.833 on the anxiety test while it was 1.5 for those who did not, illustrating a negligible relationship. Whereas the average range of IQR for those who chose was 3.1, whereas it was 3.9 for those who did not get to choose. These two contrast each other and suggest a possible correlation in variability for participants who do not choose the environment.

Our findings conclude that there is a negligible influence on change in anxiety levels due to the system environment. Due to the tendency of the average data to remain similar from the comparatives, therefore it is not possible to determine any large correlation. Additionally, people with a higher initial anxiety level tended to have the largest change. Although factors apart from system environment, and choice might have played a larger role in the resultant anxiety level due to the relatively low change found in the respective variables (Kwon, 2020). Furthermore, our findings reflect that choice had a small effect, producing narrower max, min's, and IQR's, conveying a possible relationship. Where someone's previous experience might manipulate the variability in the data due to experiences from similar environments (Matteo Girondini et al., 2023).

Overall, we came to a consensus that our findings reflect that volition might have an impact on the variability of anxiety level, where a lack of choice leads to higher variability while having a choice leads to more consistent levels. Although the changes are small they do indicate a psychological variability in the response. Whereas the actual environment will most likely not affect the change in anxiety levels.

## Professional Standard of Report

Throughout the report every member maintained respect for each others privacy, and confidentiality along with professionalism, letting each other do there own dedicated work towards our data story. Additionally, ethical principles were obeyed, with truthfulness and integrity of our work progress, keeping everyone accountable for their own actions.

------------------------------------------------------------------------

## Acknowledgements

1.  See APA information here: <https://libguides.library.usyd.edu.au/citation/apa7>
2.  The references are not included in the word count.
3.  The comments on articles can be put in a separate section, or weaved throughout the report, at appropriate places.

The report ends with an 'Acknowledgments' section. The Acknowledgment section includes:

-   a list of group meetings (date and time and attendance) - the contribution of each group member

{{< include "Meetings and Contributions.qmd" >}}

-   all resources used (eg url of stack overflow, url of Ed post, date and details of drop-in session with tutor)

{{< include "Acknowledgements and Resources.qmd" >}}

-   an AI usage statement explaining how AI was used, the name and version of AI tools used, the date/s of use of the AI tools and a link to the entire prompt session used to inform the report. If no AI was used, this must be made explicit.

## References

Kwon, C. (2020). The Effect of the Degree of Anxiety of Learners during the Use of VR on the Flow and Learning Effect. *Applied Sciences*, *10*(4932), 4932. https://doi.org/10.3390/app10144932

Matteo Girondini, Milena Stefanova, Pillan, M., & Gallace, A. (2023). *The Effect of Previous Exposure on Virtual Reality Induced Public Speaking Anxiety: A Physiological and Behavioral Study*. *26*(2), 127–133. https://doi.org/10.1089/cyber.2022.0121
