---
title: "Investigating the Effects of Choice on Mitigating Anxiety"
authors: Ari Astawa, August Gonzales, Saksham Jain, Harry Le, Leo Sharwood, Blake Walker
format:
    html:
        embed-resources: true
        toc: true
        toc-location: right
        toc-title: "Table of Contents"
        toc-depth: 3
        code-fold: true
        code-tools: true
execute:
    warning: false
    message: false
---

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
#| code-fold: true
library(tidyverse)
library(readxl)
library(gganimate)
library(gifski)
library(dplyr)
library(plotly)

# Define age groups
agegroup = Vectorize(function(number) {
	if (number >= 17 & number <= 20) {
		return ("17-20")
	} else if (number >= 21 & number <= 25) {
		return ("21-25")
	} else if (number >= 26 & number <= 30) {
		return ("26-30")
	} else if (number >= 31 & number <= 35) {
		return ("31-35")
	} else if (number >= 36 & number <= 40) {
		return ("36-40")
	} else {
		return ("41+")
    }
})

# Import data
data = read_xlsx("all_data.xlsx", sheet = "pooled123") |>
    mutate(choice = recode(choice, CHOICE = "Choice", NO_CHOICE = "No Choice")) |> 
    mutate(ageset = agegroup(age)) |> 
    filter(ageset != "41+")
# Combine choice and environment as single variable
data$choice_and_env = paste(data$choice, "-", data$env)
```

## Executive Summary/Insight

This report examines the influence of specific VR-environments on user's anxiety. We found minimal correlation between sunny and snowy environments on impacting anxiety however when given agency to choose the environment there was an observable reduction on the spread of data inclusively. Thus, providing user's choice on experienced VR-environments can potentially reduce impact on their anxiety levels.

## Initial Data Analysis

**Overview**\
The data sourced (Saunders, 2023), examines a community sample of 337 healthy adults undergoing a VR nocebo experiment. Our analysis extrapolates the data on age, anxiety levels (active and baseline) and experienced VR environments. However, there is reclassification of age into a qualitative nominal to analyse based on groups. Conversely, the VR environment remains as a qualitative nominal, anxiety level remains discrete variable, and choice remains as a qualitative normal.

**Limitations**\
The reliability of the findings are limited by the sample size and in turn become vulnerable to individual predispositions skewing results. These predispositions include sentiments resulting from geographical, cultural, societal and personal events relevant to the day of testing. Thus, to have increased reliability more control variables should be in place.

**Assumptions\
**Data cleaning was implemented to remove any outliers in regards to participant's anxiety levels exceeding the possible range as it is assumed they were mistakes. Moreover, only data on ages between 15 and 40 were examined due to insufficient participants outside this range to reach conclusive findings. Lastly, there were assumptions that: conditions between tests remained controlled to mitigate influence of confounding variables, all participants had only the permitted knowledge on how the test was conducted, all participants were honest with psychological symptoms and the recording of their anxiety levels (baseline and active), any bias participants and participants with a history of severe psychological instability were exempt from results too.

## Analysis Of Data

::: panel-tabset
## Choice & Environment vs. Δ VR Anxiety, Separation by Age Ranges

```{r}
#| code-fold: true
# Choice & Environment vs. Δ VR Anxiety separated by age ranges
ce_vranx_age_plot = ggplot(data, aes(x = choice_and_env, y = vranxiety, color = choice, fill = env)) + 
    geom_boxplot() + 
    labs(title = "Choice & Environment vs. Change in Anxiety Level by Age Ranges {closest_state}",
        x = "Choice Given & VR Environment",
        y = "Δ VR Anxiety Level",
        color = "Choice Given",
        fill = "VR Environment") +
    scale_fill_manual(values=c("lightblue", "orange")) +
    scale_colour_manual(values=c("chartreuse4", "brown3")) +
    transition_states(ageset, transition_length = 1, state_length = 1, wrap = TRUE) +
    enter_fade() + 
    exit_shrink() + 
    ease_aes('exponential-in-out')

animate(ce_vranx_age_plot, renderer = gifski_renderer(), duration = 15)
```

## Choice & Environment vs. Δ VR Anxiety, No Separation by Age Ranges

```{r}
#| code-fold: true
# Choice & Environment vs. Δ VR Anxiety separated
p =ggplot(data, aes(x = choice_and_env, y = (p_vra - BVRA), color = choice, fill = env)) + 
    geom_boxplot() + 
    labs(title = "Choice & Environment vs. Change in Anxiety Level by Age Ranges {closest_state}",
        x = "Choice Given & VR Environment",
        y = "Δ VR Anxiety Level",
        color = "Choice Given",
        fill = "VR Environment") +
    scale_fill_manual(values=c("lightblue", "orange")) +
    scale_colour_manual(values=c("chartreuse4", "brown3"))
ggplotly(p)
```

## Choice Only and Environment Only

```{r}
#| code-fold: true
# ONLY Choice vs. Δ VR Anxiety
p = ggplot(data, aes(x = choice, y = vranxiety, color = choice)) +
    labs(x = "Choice Given", y="Δ VR Anxiety", color = "Choice Given") +
    geom_boxplot() +
    scale_colour_manual(values=c("chartreuse4", "brown3"))
# ONLY Environment vs. Δ VR Anxiety
g= ggplot(data, aes(x = env, y = vranxiety, fill = env)) +
    labs(x = "VR Environment", y="Δ VR Anxiety", fill = "VR Environment") +
    geom_boxplot() +
    scale_fill_manual(values=c("lightblue", "orange"))
ggplotly(p)
ggplotly(g)
```
:::

Our analysis within this report found that there was no relationship between either the ability to choose or the VR environment to the level of anxiety change. However, it is notable that the range for change in anxiety for those experiencing a sunny environment was somewhat greater than those with a snowy environment. This is more likely to be a result of natural variation in the dataset though; there is unlikely to be an underlying mechanism for these results.

The data conveyed that the change in anxiety levels between users of sunny, and snowy environments varied by minuscule amounts. This was concluded because the median change among all age groups for sunny environments was 1.75, while snowy was 2, revealing that snowy environments had a slightly higher change. Furthermore, the average range of IQR was 4 for sunny environments, and 2.625 for snowy, implying that the IQR change was a resultant of a more decisive change.

A more detailed look at the data finds that the mean change in anxiety in participants who were given the ability to choose was 2.096 compared to 1.955 for those who did not. This demonstrates that choice has no bearing on easing the amount of anxiety a person experiences with regards to a daunting activity that may cause side effects like cyber sickness. Whereas the average range of IQR for those who chose was 3, whereas it was 4 for those who did not get to choose.

To conclude, there is a negligible influence on change in anxiety levels due to the system environment, as the average anxiety levels of different environments are very similar. Although factors apart from system environment, and choice might have played a larger role in the resultant anxiety level due to the relatively low change found in the respective variables (Kwon, 2020). Furthermore, our findings reflect that choice had a small effect, producing narrower max, min's, and IQR's, conveying a possible relationship. Where someone's previous experience might manipulate the variability in the data due to experiences from similar environments (Matteo Girondini et al., 2023). We came to a consensus that our findings reflect that volition might have an impact on the variability of anxiety level, where a lack of choice leads to higher variability while having a choice leads to more consistent levels. Although the changes are small they do indicate a psychological variability in the response. Whereas the actual environment won't impact changing anxiety levels.

## Professional Standard of Report

Throughout the report every member maintained respect for each others privacy, and confidentiality along with professionalism, letting each other do there own dedicated work towards our data story. Additionally, ethical principles were obeyed, with truthfulness and integrity of our work progress, keeping everyone accountable for their own actions.

------------------------------------------------------------------------

## References

{{< include "Meetings and Contributions.qmd" >}}

-   all resources used (eg url of stack overflow, url of Ed post, date and details of drop-in session with tutor)

{{< include "Acknowledgements and Resources.qmd" >}}

-   an AI usage statement explaining how AI was used, the name and version of AI tools used, the date/s of use of the AI tools and a link to the entire prompt session used to inform the report. If no AI was used, this must be made explicit.
